### 2024-09-24 paper/paper.qmd
First draft of analysis
Upload test scripts and changes in other scripts; list structure for the paper

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - "Angel Xu"
thanks: "Code and data are available at: https://github.com/Anjojoo/Toronto-Overnight-Servive-Occupancy-Capacity."
date: today
date-format: long
abstract: "In this report, the overnight service occupancy and capacity of Toronto is analyzed. "
format: 
  pdf: 
    toc: true
    toc-depth: 2
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(opendatatoronto)
```

# Introduction

Briefly introduce the analyzing reason and background information.
From what perspective and how to analyze
Key results
Structure introduce

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

# library(rstanarm)

# first_model <-
#   readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# modelsummary::modelsummary(
#   list(
#     "First model" = first_model
#   ),
#   statistic = "mad",
#   fmt = 2
# )
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

# pp_check(first_model) +
#   theme_classic() +
#   theme(legend.position = "bottom")
# 
# posterior_vs_prior(first_model) +
#   theme_minimal() +
#   scale_color_brewer(palette = "Set1") +
#   theme(legend.position = "bottom") +
#   coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

# plot(first_model, "trace")
# 
# plot(first_model, "rhat")
```

\newpage

# References

### 2024-09-25 paper/paper.qmd
Revision of scripts
Add and remove columns in data set

@@
 
-# Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
 # Results
@@
 
-# Discussion
-
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
-
-\newpage
-
-\appendix
-
-# Appendix {-}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-# pp_check(first_model) +
-#   theme_classic() +
-#   theme(legend.position = "bottom")
-# 
-# posterior_vs_prior(first_model) +
-#   theme_minimal() +
-#   scale_color_brewer(palette = "Set1") +
-#   theme(legend.position = "bottom") +
-#   coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-# plot(first_model, "trace")
-# 
-# plot(first_model, "rhat")
-```
-
 \newpage

### 2024-09-25 scripts/00-simulate_data.R

@@ -32,21 +32,19 @@ capacity_type <- c("Room Based Capacity", "Bed Based Capacity")
 simulated_data <- tibble(
   id = 1:number_of_datas, 
   OCCUPANCY_DATE = as.Date(runif(n = number_of_datas, min = as.numeric(start_date), 
-                                        max = as.numeric(end_date)),
-  origin = "1970-01-01"),
+                                        max = as.numeric(end_date)), 
+  origin = "1970-01-01"), 
   OVERNIGHT_SERVICE_TYPE = sample(overnight_service_type, number_of_datas, replace = TRUE), 
-  SERVICE_USER_COUNT = sample(1:660, number_of_datas, replace = TRUE), 
+  SERVICE_USER_COUNT = rpois(n = number_of_datas, lambda = 330), 
   CAPACITY_TYPE = sample(capacity_type, number_of_datas, replace = TRUE), 
   CAPACITY_ACTUAL_BED = ifelse(CAPACITY_TYPE != "Bed Based Capacity", 0, 
-                               sample(1:278, number_of_datas, replace = TRUE)), 
-  OCCUPIED_BEDS = ifelse(CAPACITY_TYPE != "Bed Based Capacity", 0, 
-                         sample(1:CAPACITY_ACTUAL_BED, number_of_datas, replace = TRUE)), 
-  OCCUPANCY_RATE_BEDS = OCCUPIED_BEDS / CAPACITY_ACTUAL_BED * 100, 
+                               rpois(n = number_of_datas, lambda = 150)), 
+  OCCUPANCY_RATE_BEDS = ifelse(CAPACITY_TYPE != "Bed Based Capacity", 0, 
+                               sample(0:100, number_of_datas, replace = TRUE)), 
   CAPACITY_ACTUAL_ROOM = ifelse(CAPACITY_TYPE == "Bed Based Capacity", 0, 
-                                sample(1:329, number_of_datas, replace = TRUE)), 
-  OCCUPIED_ROOMS = ifelse(CAPACITY_TYPE == "Bed Based Capacity", 0, 
-                          sample(1:CAPACITY_ACTUAL_ROOM, number_of_datas, replace = TRUE)), 
-  OCCUPANCY_RATE_ROOMS = OCCUPIED_ROOMS / CAPACITY_ACTUAL_ROOM * 100
+                                rpois(n = number_of_datas, lambda = 170)), 
[…]

### 2024-09-25 scripts/02-data_cleaning.R

@@ -9,23 +9,29 @@
 #### Workspace setup ####
 library(tidyverse)
 library(janitor)
+library(lubridate)
+library(dplyr)
 
 #### Clean data ####
 # Read the raw dataset
 raw_data <- read_csv("data/raw_data/raw_data.csv")
 
-# Clean and select wanted columns
+# Convert occupancy_date into date format
+raw_data$OCCUPANCY_DATE <- as.Date(raw_data$OCCUPANCY_DATE, format = "%y-%m-%d")
+
+# Clean, mutate and select wanted columns
 cleaned_data <-
   raw_data |>
-  janitor::clean_names() |>
-  select(overnight_service_type, 
+  janitor::clean_names() |> 
+  mutate(year = year(occupancy_date)) |> 
+  select(year, 
+         occupancy_date, 
+         overnight_service_type, 
          service_user_count, 
          capacity_type, 
          capacity_actual_bed,
-         occupied_beds,
          occupancy_rate_beds,
          capacity_actual_room,
-         occupied_rooms,
          occupancy_rate_rooms)
 
 # View the cleaned dataset (optional)

### 2024-09-25 scripts/03-test_data.R

@@ -11,13 +11,18 @@
 library(tidyverse)
 
 #### Test data ####
-data <- read_csv("data/raw_data/raw_data.csv")
+data <- read_csv("data/analysis_data/analysis_data.csv")
 
 # Test for negative numbers in numerical columns
-data$SERVICE_USER_COUNT |> min() >= 0
+data$service_user_count |> min() >= 0
+any_negative_bed<- any(data$capacity_actual_bed |> min() >= 0, na.rm = TRUE)
+any_negative_room<- any(data$capacity_actual_room |> min() >= 0, na.rm = TRUE)
+any_negative_rate_bed<- any(data$occupancy_rate_beds |> min() >= 0, na.rm = TRUE)
+any_negative_rate_room<- any(data$occupancy_rate_rooms |> min() >= 0, na.rm = TRUE)
+all(any_negative_bed, any_negative_room, any_negative_rate_bed, any_negative_rate_room)
 
 # Test for NA in categorial columns
-all(!is.na(data$OVERNIGHT_SERVICE_TYPE))
-all(!is.na(data$CAPACITY_TYPE))
+all(!is.na(data$overnight_service_type))
+all(!is.na(data$capacity_type))
 
 

### 2024-09-25 scripts/00-simulate_data.R
Revision of scripts

@@ -32,7 +32,6 @@ capacity_type <- c("Room Based Capacity", "Bed Based Capacity")
 simulated_data <- tibble(
   id = 1:number_of_datas, 
   OCCUPANCY_DATE = as.Date(runif(n = number_of_datas, min = as.numeric(start_date), 
-<<<<<<< HEAD
                                         max = as.numeric(end_date)), 
   origin = "1970-01-01"), 
   OVERNIGHT_SERVICE_TYPE = sample(overnight_service_type, number_of_datas, replace = TRUE), 
@@ -46,23 +45,6 @@ simulated_data <- tibble(
                                 rpois(n = number_of_datas, lambda = 170)), 
   OCCUPANCY_RATE_ROOMS = ifelse(CAPACITY_TYPE != "Bed Based Capacity", 0, 
                                 sample(0:100, number_of_datas, replace = TRUE))
-=======
-                                        max = as.numeric(end_date)),
-  origin = "1970-01-01"),
-  OVERNIGHT_SERVICE_TYPE = sample(overnight_service_type, number_of_datas, replace = TRUE), 
-  SERVICE_USER_COUNT = sample(1:660, number_of_datas, replace = TRUE), 
-  CAPACITY_TYPE = sample(capacity_type, number_of_datas, replace = TRUE), 
-  CAPACITY_ACTUAL_BED = ifelse(CAPACITY_TYPE != "Bed Based Capacity", 0, 
-                               sample(1:278, number_of_datas, replace = TRUE)), 
-  OCCUPIED_BEDS = ifelse(CAPACITY_TYPE != "Bed Based Capacity", 0, 
-                         sample(1:CAPACITY_ACTUAL_BED, number_of_datas, replace = TRUE)), 
-  OCCUPANCY_RATE_BEDS = OCCUPIED_BEDS / CAPACITY_ACTUAL_BED * 100, 
-  CAPACITY_ACTUAL_ROOM = ifelse(CAPACITY_TYPE == "Bed Based Capacity", 0, 
-                                sample(1:329, number_of_datas, replace = TRUE)), 
-  OCCUPIED_ROOMS = ifelse(CAPACITY_TYPE == "Bed Based Capacity", 0, 
-                          sample(1:CAPACITY_ACTUAL_ROOM, number_of_datas, replace = TRUE)), 
-  OCCUPANCY_RATE_ROOMS = OCCUPIED_ROOMS / CAPACITY_ACTUAL_ROOM * 100
->>>>>>> origin/main
 )
 
 #### Write_csv ####

### 2024-09-25 scripts/02-data_cleaning.R

@@ -9,17 +9,14 @@
 #### Workspace setup ####
 library(tidyverse)
 library(janitor)
-<<<<<<< HEAD
 library(lubridate)
 library(dplyr)
-=======
->>>>>>> origin/main
+
 
 #### Clean data ####
 # Read the raw dataset
 raw_data <- read_csv("data/raw_data/raw_data.csv")
 
-<<<<<<< HEAD
 # Convert occupancy_date into date format
 raw_data$OCCUPANCY_DATE <- as.Date(raw_data$OCCUPANCY_DATE, format = "%y-%m-%d")
 
@@ -36,20 +33,6 @@ cleaned_data <-
          capacity_actual_bed,
          occupancy_rate_beds,
          capacity_actual_room,
-=======
-# Clean and select wanted columns
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(overnight_service_type, 
-         service_user_count, 
-         capacity_type, 
-         capacity_actual_bed,
-         occupied_beds,
-         occupancy_rate_beds,
-         capacity_actual_room,
-         occupied_rooms,
->>>>>>> origin/main
          occupancy_rate_rooms)
 
 # View the cleaned dataset (optional)

### 2024-09-25 scripts/03-test_data.R

@@ -11,7 +11,6 @@
 library(tidyverse)
 
 #### Test data ####
-<<<<<<< HEAD
 data <- read_csv("data/analysis_data/analysis_data.csv")
 
 # Test for negative numbers in numerical columns
@@ -25,15 +24,5 @@ all(any_negative_bed, any_negative_room, any_negative_rate_bed, any_negative_rat
 # Test for NA in categorial columns
 all(!is.na(data$overnight_service_type))
 all(!is.na(data$capacity_type))
-=======
-data <- read_csv("data/raw_data/raw_data.csv")
-
-# Test for negative numbers in numerical columns
-data$SERVICE_USER_COUNT |> min() >= 0
-
-# Test for NA in categorial columns
-all(!is.na(data$OVERNIGHT_SERVICE_TYPE))
-all(!is.na(data$CAPACITY_TYPE))
->>>>>>> origin/main
 
 

### 2024-09-25 paper/paper.qmd
Sketches update and remove useless sections in paper

@@
 
-# Results
-
-Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-# library(rstanarm)
-
-# first_model <-
-#   readRDS(file = here::here("models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-# modelsummary::modelsummary(
-#   list(
-#     "First model" = first_model
-#   ),
-#   statistic = "mad",
-#   fmt = 2
-# )
-```
-
 \newpage

### 2024-09-25 paper/paper.qmd
Update LLM usage and add reference

@@
 library(opendatatoronto)
+library(here)
+
+# Load the analysis data
+data <- read_csv(here::here("data", "analysis_data", "analysis_data.csv"))
 ```
@@
 #| echo: false
+#| warning: false
+#| message: false
 

### 2024-09-26 paper/paper.qmd
Paper draft 2_Data
Adding data parts into the paper

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "An Analysis of daily shelter and overnight service usage in Toronto focusing on shelter occupancy and capacity"
 author: 
@@
 library(here)
+library(ggplot2)
+library(knitr)
 
@@
 
-You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
+You can and should cross-reference sections and sub-sections. We use @citeR and @alexander2023telling.
 
@@
 # Data {#sec-data}
+## Data Source and Overview
+This paper analyzes data from the "Daily Shelter & Overnight Service Occupancy & Capacity" dataset available on the OpenDataToronto portal.
+The dataset contains a daily record of overnight shelters and service programs collected through Toronto’s Shelter Management Information System (SMIS).
+Data spans from 2021 to 2024, offering current insights, and is updated daily, except for weekends and holidays, with uploads resuming on the next business day.
 
+A similar dataset, "Daily Shelter Occupancy," is also available on OpenDataToronto but  only provides data up to March 10, 2022, and includes fewer variables.
+In contrast, "Daily Shelter & Overnight Service Occupancy & Capacity" features the most recent data, refreshed on 25 September, 2024.
+It builds upon and enhances the earlier dataset by incorporating additional key variables, including "overnight service type" and "actual capacity." The "overnight service type" variable categorizes the types of overnight services offered in Toronto's shelters, helping analyze variations in user counts by service type.
+The "actual capacity" variable represents the true available capacity of shelters, as opposed to the funding-based capacity in the previous dataset, offering a more accurate measure of occupancy rates for further analysis.
+
+## Data Tools
+The data was simulated, extracted, and cleaned by R[@citeR] from the opendatatoronto package[@citeopendatatoronto].
+Tidyverse[@citetidyverse], here[@citehere], ggplot2[@citeggplot2], dplyr[@citedplyr], lubridate[@lubridate], janitor[@citejanitor] and knitr[@knitr] are all applied in the analysis.
+
+## Clean Data
+Prior to analysis, data cleaning is essential to remove unnecessary columns and create new ones for further examination.
+The occupancy date variable is converted to a date format, allowing the extraction of the year for subsequent analysis.
+After this conversion, a new 'year' column is mutated to show trends in service user counts and occupancy rates from 2021 to 2024.
+Once these steps are complete, relevant variables are selected from the raw dataset to simplify and clarify the analysis.
+The chosen variables include year, occupancy date, overnight service type, service user count, capacity type, actual capacity and occupancy rates.
+NAs exist in actual capacity and occupancy rates, but there are no needs to eliminate them since due to different capacity type of bed and room, NA appears in the opposite actual capacity and occupancy rates.
+
+## Summary Statistics of Variables
+To gain a better overview of the dataset, a table including key variables is provided. Table 1(@tbl-key-variables) presents year, overnight service type, service user count and occupancy rates.
+Year represents the year where the observation is collected.
+Overnight service type represents what service has been provided.
+Service user count records number of users taking service during that night.
+Occupancy rates record the proportion of actual capacity that is occupied.
+
+```{r}
+#| label: tbl-key-variables
+#| echo: false
+#| warning: false
+#| message: false
+
+#  Create a subset includes key variables
+subset <- data[, c("year", "overnight_service_type", "service_user_count", "occupancy_rate_beds", 
+                   "occupancy_rate_rooms")]
+
+# Present the first ten rows of key variables as a sample
+sample <- head(subset)
+
+# Present the sample as a table
+kable(sample, col.names = c("year", "overnight service type", "service user count", 
+                            "occupancy rate beds", "occupancy rate rooms"), 
+      align = "ccc", 
+      caption = "Table 1: Sample of the key variables in Occupancy & Capacity Data")
+```
+
+Beside the sample table of ley variables, a summary statistics table of key variables can be provided to clarify some information of the data.
+Table 2(@tbl-summary-stats) presents the mean and standard deviation of service user counts and two occupancy rates.
+Service user counts has a mean at 73.73 and a standard deviation at about 91.68% which is pretty high and shows the great difference in user picks between different services.
+Occupancy rate of beds has a high mean at 98.24% and a standard deviation at 6.08%.
+This illustrates that the occupency rate of beds are approximately near 100 and are full occupied for most of the situations.
+Occupancy rate of rooms also has a high mean at 99.53% which is higher than occupancy rate of beds and a even lower standard deviation of 1.77%, suggesting that capacity of shelter rooms reaches nearly full and fuller than shelter rooms.
+However, the actual capacity of beds and rooms aren't included in this table; thus, the usage of shelter beds and rooms need to be considered more thoroughly by observing their actual capacity numbers.
+
+```{r}
+#| label: tbl-summary-stats
+#| echo: false
+#| warning: false
+#| message: false
+
+# Calculate the mean of variables
+mean_user_cnt <- mean(data$service_user_count)
+mean_rate_bed <- mean(data$occupancy_rate_beds, na.rm = TRUE)
+mean_rate_room <- mean(data$occupancy_rate_rooms, na.rm = TRUE)
+
+# Calculate the sd of variables
+sd_user_cnt <- sd(data$service_user_count)
+sd_rate_bed <- sd(data$occupancy_rate_beds, na.rm = TRUE)
+sd_rate_room <- sd(data$occupancy_rate_rooms, na.rm = TRUE)
+
+#  Create a subset includes key variables
+summary_stats <- data.frame(Variable = c("Service User Count", "Occupancy Rate of Beds", 
+                                         "Occupancy Rate of Rooms"), 
+                            Mean = c(mean_user_cnt, mean_rate_bed, mean_rate_room), 
+                            SD = c(sd_user_cnt, sd_rate_bed, sd_rate_room))
+
+# Present the sample as a table
+kable(summary_stats, col.names = c("Variable", "Mean", "Standard Deviation"), 
+      align = "ccc", 
+      caption = "Table 2: Mean and Standard Deviation of Key Variables")
+```
+
+# Result
 Some of our data is of penguins (@fig-bills), from @palmerpenguins.
@@
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: fig-service-types
+#| fig-cap: Total service user counts for different overnight service types
 #| echo: false
@@
 #| message: false
+
+data |> ggplot(aes(x = overnight_service_type, y = service_user_count)) +  
+  geom_col() + 
+  theme_minimal() + 
+  labs(x = "overnight service types", y = "service user counts") + 
+  coord_flip()
 
@@
 ```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
+#| label: fig-year-count
+#| fig-cap: Service user count change from 2021 to 2024
 #| echo: false
@@
 
-analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
+data |> ggplot(aes(x = year, y = service_user_count)) +  
+  geom_col() + 
+  theme_minimal() + 
+  labs(x = "years", y = "service user counts") + 
+  coord_flip()
 

### 2024-09-26 paper/paper.qmd
Update Sketches

@@
 library(knitr)
+library(dplyr)
 
@@
 Table 2(@tbl-summary-stats) presents the mean and standard deviation of service user counts and two occupancy rates.
-Service user counts has a mean at 73.73 and a standard deviation at about 91.68% which is pretty high and shows the great difference in user picks between different services.
-Occupancy rate of beds has a high mean at 98.24% and a standard deviation at 6.08%.
+Service user counts has a mean at 60.58 and a standard deviation at about 76.09% which is pretty high and shows the great difference in user picks between different services.
+Occupancy rate of beds has a high mean at 96.24% and a standard deviation at 9.28%.
 This illustrates that the occupency rate of beds are approximately near 100 and are full occupied for most of the situations.
-Occupancy rate of rooms also has a high mean at 99.53% which is higher than occupancy rate of beds and a even lower standard deviation of 1.77%, suggesting that capacity of shelter rooms reaches nearly full and fuller than shelter rooms.
+Occupancy rate of rooms also has a high mean at 97.03% which is slightly higher than occupancy rate of beds and a standard deviation of 10.82%, suggesting that capacity of shelter rooms reaches nearly full and fuller than shelter rooms.
 However, the actual capacity of beds and rooms aren't included in this table; thus, the usage of shelter beds and rooms need to be considered more thoroughly by observing their actual capacity numbers.
@@
       caption = "Table 2: Mean and Standard Deviation of Key Variables")
+```
+
+For deeper explore in service user count and capacity in different years, the sum of these variables from 2021 to 2024 needs to be examined.
+Occupancy rate of beds and rooms cannot reflect the total difference usage of all shelters, as well as the mean of service user counts.
+They can only represent the common usage situation of every shelter in total.
+While Table 3(@tbl-total) provides a table of total service user counts and actual capacity for both beds and rooms.
+Within these data, a more complete result can be figured out and will be used for data visualization later.
+
+```{r}
+#| label: tbl-total
+#| echo: false
+#| warning: false
+#| message: false
+
+# Calculate the sum of service user counts in each year
+sum_cnt_2021 <- sum(data$service_user_count[data$year == 2021], na.rm = TRUE)
+sum_cnt_2022 <- sum(data$service_user_count[data$year == 2022], na.rm = TRUE)
+sum_cnt_2023 <- sum(data$service_user_count[data$year == 2023], na.rm = TRUE)
+sum_cnt_2024 <- sum(data$service_user_count[data$year == 2024], na.rm = TRUE)
+
+# Calculate the sum of actual bed capacity in each year
+sum_bed_2021 <- sum(data$capacity_actual_bed[data$year == 2021], na.rm = TRUE)
+sum_bed_2022 <- sum(data$capacity_actual_bed[data$year == 2022], na.rm = TRUE)
+sum_bed_2023 <- sum(data$capacity_actual_bed[data$year == 2023], na.rm = TRUE)
+sum_bed_2024 <- sum(data$capacity_actual_bed[data$year == 2024], na.rm = TRUE)
+
+# Calculate the sum of actual bed capacity in each year
+sum_room_2021 <- sum(data$capacity_actual_room[data$year == 2021], na.rm = TRUE)
+sum_room_2022 <- sum(data$capacity_actual_room[data$year == 2022], na.rm = TRUE)
+sum_room_2023 <- sum(data$capacity_actual_room[data$year == 2023], na.rm = TRUE)
+sum_room_2024 <- sum(data$capacity_actual_room[data$year == 2024], na.rm = TRUE)
+
+#  Create a subset includes key variables
+summary_stats <- data.frame(Year = c("2021", "2022", "2023", "2024"), 
+                            Service_User_Counts_Sum = c(sum_cnt_2021, sum_cnt_2022, 
+                                                       sum_cnt_2023, sum_cnt_2024), 
+                            Actual_Bed_Capacity_Sum = c(sum_bed_2021, sum_bed_2022, 
+                                                        sum_bed_2023, sum_bed_2024), 
+                            Actual_Room_Capacity_Sum = c(sum_room_2021, sum_room_2022, 
+                                                         sum_room_2023, sum_room_2024)
+                            )
+
+# Present the sample as a table
+kable(summary_stats, col.names = c("Year", "Total Service User Counts", 
+                                   "Total Actual Bed Capacity", "Total Actual Room Capacity"), 
+      align = "ccc", 
+      caption = "Table 3: Total Service User Count and Actual Capacity between years")
 ```
@@
 
+data$year <- as.factor(data$year)
+
 data |> ggplot(aes(x = year, y = service_user_count)) +  
@@
   labs(x = "years", y = "service user counts") +  
-  coord_flip()
-
+  scale_y_continuous(labels = scales::comma)
 ```
@@
 
+```{r}
+#| label: fig-year-count
+#| fig-cap: Service user count change from 2021 to 2024
+#| echo: false
+#| warning: false
+#| message: false
+
+data$year <- as.factor(data$year)
+
+data |> ggplot(aes(x = year, y = occurency_rate_beds)) +  
+  geom_line() + 
+  theme_minimal() + 
+  labs(x = "years", y = "service user counts") + 
+  scale_y_continuous(labels = scales::comma)
+
+```
+
 \newpage

### 2024-09-26 scripts/02-data_cleaning.R

@@ -14,11 +14,20 @@ library(dplyr)
 
 
 #### Clean data ####
-# Read the raw dataset
-raw_data <- read_csv("data/raw_data/raw_data.csv")
+# Read 2021-2024 raw datasets
+raw_data_2021 <- read_csv("data/raw_data/raw_data_2021.csv")
+raw_data_2022 <- read_csv("data/raw_data/raw_data_2022.csv")
+raw_data_2023 <- read_csv("data/raw_data/raw_data_2023.csv")
+raw_data_2024 <- read_csv("data/raw_data/raw_data_2024.csv")
 
 # Convert occupancy_date into date format
-raw_data$OCCUPANCY_DATE <- as.Date(raw_data$OCCUPANCY_DATE, format = "%y-%m-%d")
+raw_data_2021$OCCUPANCY_DATE <- as.Date(raw_data_2021$OCCUPANCY_DATE, format = "%y-%m-%d")
+raw_data_2022$OCCUPANCY_DATE <- as.Date(raw_data_2022$OCCUPANCY_DATE, format = "%y-%m-%d")
+raw_data_2023$OCCUPANCY_DATE<- as.Date(raw_data_2023$OCCUPANCY_DATE, format="%Y-%m-%dT%H:%M:%S")
+raw_data_2024$OCCUPANCY_DATE <- as.Date(raw_data_2024$OCCUPANCY_DATE, format = "%y-%m-%d")
+
+# Combine 2021-2024 datasets into one dataset
+raw_data <- rbind(raw_data_2021, raw_data_2022, raw_data_2023, raw_data_2024)
 
 # Clean, mutate and select wanted columns
 cleaned_data <-
@@ -33,7 +42,8 @@ cleaned_data <-
          capacity_actual_bed,
          occupancy_rate_beds,
          capacity_actual_room,
-         occupancy_rate_rooms)
+         occupancy_rate_rooms) |> 
+  drop_na(overnight_service_type)
 
 # View the cleaned dataset (optional)
 head(cleaned_data)

### 2024-09-26 scripts/03-test_data.R

@@ -24,5 +24,6 @@ all(any_negative_bed, any_negative_room, any_negative_rate_bed, any_negative_rat
 # Test for NA in categorial columns
 all(!is.na(data$overnight_service_type))
 all(!is.na(data$capacity_type))
+all(!is.na(data$year))
 
 

### 2024-09-26 paper/paper.qmd
Add tests for simulated data

@@
 date-format: long
-abstract: "In this report, the overnight service occupancy and capacity of Toronto is analyzed. "
+abstract: "This paper analyzes the overnight service occupancy and capacity of shelters in Toronto, utilizing data from OpenDataToronto.
+Employing the R programming language along with various packages, the analysis focuses on shelter usage trends, specifically examining changes in occupancy rates and actual capacity over the years.
+The results indicate an overall upward trend in shelter occupancy rates, while a significant decline in actual capacity was observed in 2024.
+These findings highlight the need for strategic resource allocation and improved response strategies concerning shelter occupancy in the future."
 format: 
@@
 
-Briefly introduce the analyzing reason and background information.
-From what perspective and how to analyze
-Key results
-Structure introduce
+The growing issue of homelessness in urban centers like Toronto has created an urgent need to analyze shelter occupancy and capacity, as shelters are critical for providing temporary housing to those in need.
+To effectively manage and improve these services, understanding usage trends is essential.
+This paper focuses on analyzing the "Daily Shelter & Overnight Service Occupancy & Capacity" dataset from OpenDataToronto(@citeopendatatoronto), covering data from 2021 to 2024.
+This dataset includes daily records of shelter occupancy, capacity, and types of overnight services provided, offering a comprehensive view of the situation in Toronto.
+The motivation behind this study is to address the fluctuating demand for shelter services and inform strategies for more efficient allocation of resources to support vulnerable populations.
 
-You can and should cross-reference sections and sub-sections. We use @citeR and @alexander2023telling.
+The study involves a detailed analysis of shelter usage trends, employing R programming(@citeR) and @alexander2023telling mainly.
+Besides, associated packages of R are applied to clean, manipulate, and visualize the data.
+Tools like tidyverse(@citetidyverse), ggplot2(@citeggplot2), dplyr(@citedplyr), and others facilitated data processing and visualization, enabling a deeper examination of occupancy rates and capacity changes over the years.
+The findings reveal an overall upward trend in shelter occupancy rates, suggesting increased demand for these services.
+A significant decline in actual capacity was observed in 2024, which underscores a mismatch between available resources and the needs of shelter users.
+Summary statistics were computed to highlight trends in service user counts and occupancy rates, providing insights into how shelter usage has evolved.
 
-The remainder of this paper is structured as follows. @sec-data....
+This paper's findings are crucial for policy makers, social service organizations, and urban planners.
+By shedding light on the trends in shelter occupancy and capacity, the study identifies gaps in service and potential areas for improvement.
+The analysis has implications for future resource allocation, as understanding the patterns of shelter usage can help design better support strategies, optimize capacity, and improve the overall response to homelessness in urban areas.
+Additionally, it can serve as a model for analyzing shelter systems in other cities, offering a framework for data-driven decision-making.
+
+The remainder of this paper is structured as follows.
+Section 2(@sec-data) provides details on the data source, including the specific variables used, and describes the data cleaning process to prepare for analysis.
+This data section also outlines the tools and methodologies applied.
+Section 3(@sec-results) presents the key results, focusing on occupancy rates, service user counts, and actual capacity over time.
+Visualizations are used to intuitively convey these findings.
+Section 4{@sec-dis} offers a discussion of the results and their implications, summarizing conclusions and emphasizing the importance of balancing shelter occupancy and capacity for effective service provision.
+The comprehensive structure ensures that the reader gains a full understanding of what was analyzed, the findings, and their significance.
 
@@
 The data was simulated, extracted, and cleaned by R[@citeR] from the opendatatoronto package[@citeopendatatoronto].
-Tidyverse[@citetidyverse], here[@citehere], ggplot2[@citeggplot2], dplyr[@citedplyr], lubridate[@lubridate], janitor[@citejanitor] and knitr[@knitr] are all applied in the analysis.
+Tidyverse[@citetidyverse], here[@citehere], ggplot2[@citeggplot2], dplyr[@citedplyr], lubridate[@citelubridate], janitor[@citejanitor] and knitr[@citeknitr] are all applied in the analysis.
+
+## Data Measurement and Ethics
 
@@
 
-# Result
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+# Result{#sec-results}
+Data visualizations can be used to present results more intuitively, and the data used to visualize  is from @citeopendatatoronto.
+
+## Overnight Service Usage
+The graph below displays the user counts for various overnight service types, offering a clear comparison. This visualization helps identify the most frequently used service across all types.
 
@@
 data |> ggplot(aes(x = overnight_service_type, y = service_user_count)) +  
-  geom_col() + 
+  geom_col(fill = "orange") + 
   theme_minimal() + 
   labs(x = "overnight service types", y = "service user counts") + 
-  coord_flip()
+  coord_flip() + 
+  scale_y_continuous(labels = scales::comma)
 
@@
 
-Talk more about it.
+@fig-service-types shows the distribution of overnight service usage.
+It is evident that Shelter and Motel/Hotel Shelter services have the highest user counts, with values nearly identical.
+The 24-Hour Respite Site ranks third, while the remaining services have significantly fewer users compared to these top three.
+This highlights that Shelter and Motel/Hotel Shelter are the most commonly chosen services in Toronto, likely due to the more comprehensive support they offer compared to other service types.
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+## Service Count Change
 
@@
 data |> ggplot(aes(x = year, y = service_user_count)) +  
-  geom_col() + 
+  geom_col(color = "orange") + 
   theme_minimal() + 
@@
 
-Talk way more about it.
+Figure 2(@fig-year-count) shows the number of service users across different years from 2021 to 2024, as also detailed in Table 3(@tbl-total).
+The graph highlights a clear upward trend from 2021 to 2023, followed by a decline in 2024.
+However, the total number of service users in 2024 remains higher than in 2021.
+
+## Average Occupancy Rate Change
 
 ```{r}
-#| label: fig-year-count
-#| fig-cap: Service user count change from 2021 to 2024
+#| label: fig-rate-mean
+#| fig-cap: Average Occupancy Rate of Beds and Rooms from 2021 to 2024
 #| echo: false
@@
 
-data$year <- as.factor(data$year)
+avg_rate <- data %>%
+  group_by(year) %>%
+  summarize(
+    avg_rate_bed = mean(occupancy_rate_beds, na.rm = TRUE),
+    avg_rate_room = mean(occupancy_rate_rooms, na.rm = TRUE)
+  )
 
-data |> ggplot(aes(x = year, y = occurency_rate_beds)) +  
-  geom_line() + 
+# Reshape the data to long format for plotting with side-by-side bars
+df_long <- tidyr::pivot_longer(avg_rate, cols = c("avg_rate_bed", "avg_rate_room"),
+                               names_to = "type",
+                               values_to = "average")
+
+# Plot using ggplot2
+ggplot(df_long, aes(x = factor(year), y = average, fill = type)) +
+  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) + 
+  labs(x = "Year", y = "Average Occupancy Rate", fill = "Type") +
   theme_minimal() + 
-  labs(x = "years", y = "service user counts") + 
-  scale_y_continuous(labels = scales::comma)
+  geom_text(aes(label = round(average, 1)), 
+            position = position_dodge(width = 0.9), 
+            vjust = -0.5, size = 3.5)
 
 ```
+
+Figure 3(@fig-rate-mean) illustrates the trend in the average occupancy rates of beds and rooms from 2021 to 2024.
+Both categories exhibit an upward trend, although the increase in the occupancy rates of beds and rooms from year to year is quite modest.
+Overall, the average occupancy rate for rooms consistently surpasses that of beds each year.
+In 2024, the average room occupancy rate reached 99.5%, indicating nearly full occupancy of shelter rooms, while the average bed occupancy rate achieved 98.3%, slightly lower than the average room occupancy rate.
+
+## Total Actual Capacity Change
+
+```{r}
+#| label: fig-total-cap
+#| fig-cap: Total Actual Capacity of Beds and Rooms from 2021 to 2024
+#| echo: false
+#| warning: false
+#| message: false
+
+total_cap <- data %>%
+  group_by(year) %>%
+  summarize(
+    total_cap_bed = sum(capacity_actual_bed, na.rm = TRUE),
+    total_cap_room = sum(capacity_actual_room, na.rm = TRUE)
+  )
+
+# Reshape the data to long format for plotting with side-by-side bars
+df_long <- tidyr::pivot_longer(total_cap, cols = c("total_cap_bed", "total_cap_room"),
+                               names_to = "type",
+                               values_to = "total")
+
+# Plot using ggplot2
+ggplot(df_long, aes(x = factor(year), y = total, fill = type)) +
+  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) + 
+  labs(x = "Year", y = "Total Actual Capacity", fill = "Type") +
+  theme_minimal()
+
+```
+
+Figure 4(@fig-total-cap) illustrates the total actual capacity of beds and rooms from 2021 to 2024, allowing for a comparison of the capacities across two categories.
+Notably, as shown in Figure 3(@fig-rate-mean), rooms consistently exhibit a higher occupancy rate than beds over all four years.
+The plots indicate that the total actual capacity of rooms was only greater than that of beds in 2021 and 2022; however, in 2023 and 2024, the capacity of beds surpassed that of rooms significantly.
+Additionally, the total actual capacity of beds demonstrated an increasing trend from 2021 to 2023, followed by a decline in 2024.
+Conversely, the total actual capacity of rooms exhibited an upward trend from 2021 to 2022, after which it began to decrease until reaching its lowest point in 2024.
+
+# Discussion {#sec-dis}
+## Summary Conclusions
 

### 2024-09-26 scripts/00-simulate_data.R

@@ -9,6 +9,7 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(lubridate)
 
 #### Simulate data ####
 set.seed(102)
@@ -34,6 +35,7 @@ simulated_data <- tibble(
   OCCUPANCY_DATE = as.Date(runif(n = number_of_datas, min = as.numeric(start_date), 
                                         max = as.numeric(end_date)), 
   origin = "1970-01-01"), 
+  Year = year(OCCUPANCY_DATE), 
   OVERNIGHT_SERVICE_TYPE = sample(overnight_service_type, number_of_datas, replace = TRUE), 
   SERVICE_USER_COUNT = rpois(n = number_of_datas, lambda = 330), 
   CAPACITY_TYPE = sample(capacity_type, number_of_datas, replace = TRUE), 

### 2024-09-26 scripts/03-test_data.R

@@ -10,20 +10,41 @@
 #### Workspace setup ####
 library(tidyverse)
 
-#### Test data ####
+#### Test simulated data ####
+simulated_data <- read_csv("data/raw_data/simulated_data.csv")
+
+# Test for negative numbers in numerical columns
+simulated_data$SERVICE_USER_COUNT |> min() >= 0
+simulated_any_negative_bed <- any(simulated_data$CAPACITY_ACTUAL_BED |> min() < 0, 
+                                  na.rm = TRUE)
+simulated_any_negative_room <- any(simulated_data$CAPACITY_ACTUAL_ROOM |> min() < 0, 
+                                   na.rm = TRUE)
+simulated_any_negative_rate_bed <- any(simulated_data$OCCUPANCY_RATE_BEDS |> min() < 0, 
+                                       na.rm = TRUE)
+simulated_any_negative_rate_room <- any(simulated_data$OCCUPANCY_RATE_ROOMS |> min() < 0, 
+                                        na.rm = TRUE)
+all(simulated_any_negative_bed, simulated_any_negative_room, 
+    simulated_any_negative_rate_bed, simulated_any_negative_rate_room)
+
+# Test for NA in categorical columns
+all(is.na(simulated_data$OVERNIGHT_SERVICE_TYPE))
+all(is.na(simulated_data$CAPACITY_TYPE))
+all(is.na(simulated_data$Year))
+
+#### Test cleaned data ####
 data <- read_csv("data/analysis_data/analysis_data.csv")
 
 # Test for negative numbers in numerical columns
 data$service_user_count |> min() >= 0
-any_negative_bed<- any(data$capacity_actual_bed |> min() >= 0, na.rm = TRUE)
-any_negative_room<- any(data$capacity_actual_room |> min() >= 0, na.rm = TRUE)
-any_negative_rate_bed<- any(data$occupancy_rate_beds |> min() >= 0, na.rm = TRUE)
-any_negative_rate_room<- any(data$occupancy_rate_rooms |> min() >= 0, na.rm = TRUE)
+any_negative_bed <- any(data$capacity_actual_bed |> min() < 0, na.rm = TRUE)
+any_negative_room <- any(data$capacity_actual_room |> min() < 0, na.rm = TRUE)
+any_negative_rate_bed <- any(data$occupancy_rate_beds |> min() < 0, na.rm = TRUE)
[…]

### 2024-09-26 paper/paper.qmd
Finish discussion part in paper

@@
 # Introduction
-
 The growing issue of homelessness in urban centers like Toronto has created an urgent need to analyze shelter occupancy and capacity, as shelters are critical for providing temporary housing to those in need.
@@
 ## Summary Statistics of Variables
-To gain a better overview of the dataset, a table including key variables is provided. Table 1(@tbl-key-variables) presents year, overnight service type, service user count and occupancy rates.
+To gain a better overview of the dataset, a table including key variables is provided. @tbl-key-variables presents year, overnight service type, service user count and occupancy rates.
 Year represents the year where the observation is collected.
@@
 Beside the sample table of ley variables, a summary statistics table of key variables can be provided to clarify some information of the data.
-Table 2(@tbl-summary-stats) presents the mean and standard deviation of service user counts and two occupancy rates.
+@tbl-summary-stats presents the mean and standard deviation of service user counts and two occupancy rates.
 Service user counts has a mean at 60.58 and a standard deviation at about 76.09% which is pretty high and shows the great difference in user picks between different services.
@@
 They can only represent the common usage situation of every shelter in total.
-While Table 3(@tbl-total) provides a table of total service user counts and actual capacity for both beds and rooms.
+While @tbl-total provides a table of total service user counts and actual capacity for both beds and rooms.
 Within these data, a more complete result can be figured out and will be used for data visualization later.
@@
 # Result{#sec-results}
-Data visualizations can be used to present results more intuitively, and the data used to visualize  is from @citeopendatatoronto.
+Data visualizations can be used to present results more intuitively, and the data used to visualize  is from opendatatoronto@citeopendatatoronto.
 
@@
 
-Figure 2(@fig-year-count) shows the number of service users across different years from 2021 to 2024, as also detailed in Table 3(@tbl-total).
-The graph highlights a clear upward trend from 2021 to 2023, followed by a decline in 2024.
+@fig-year-count shows the number of service users across different years from 2021 to 2024, as also detailed in @tbl-total.
+The graph highlights a clear upward trend from 2021 to 2023 and reached to a peak of 3288505 in 2023, followed by a decline back to 2633817 in 2024.
 However, the total number of service users in 2024 remains higher than in 2021.
@@
 
-Figure 3(@fig-rate-mean) illustrates the trend in the average occupancy rates of beds and rooms from 2021 to 2024.
+@fig-rate-mean illustrates the trend in the average occupancy rates of beds and rooms from 2021 to 2024.
 Both categories exhibit an upward trend, although the increase in the occupancy rates of beds and rooms from year to year is quite modest.
@@
 
-Figure 4(@fig-total-cap) illustrates the total actual capacity of beds and rooms from 2021 to 2024, allowing for a comparison of the capacities across two categories.
-Notably, as shown in Figure 3(@fig-rate-mean), rooms consistently exhibit a higher occupancy rate than beds over all four years.
-The plots indicate that the total actual capacity of rooms was only greater than that of beds in 2021 and 2022; however, in 2023 and 2024, the capacity of beds surpassed that of rooms significantly.
-Additionally, the total actual capacity of beds demonstrated an increasing trend from 2021 to 2023, followed by a decline in 2024.
-Conversely, the total actual capacity of rooms exhibited an upward trend from 2021 to 2022, after which it began to decrease until reaching its lowest point in 2024.
+@fig-total-cap illustrates the total actual capacity of beds and rooms from 2021 to 2024, allowing for a comparison of the capacities across two categories.
+Notably, as shown in @fig-rate-mean, rooms consistently exhibit a higher occupancy rate than beds over all four years.
+The plots indicate that the total actual capacity of rooms was only greater than that of beds in 2021 and 2022; however, in 2023 and 2024, the capacity of beds surpassed that of rooms significantly and raised to 1457984.
+Additionally, the total actual capacity of beds demonstrated an increasing trend from 2021 to 2023, followed by a decline to 1232028 in 2024.
+Conversely, the total actual capacity of rooms exhibited an upward trend from 2021 to 2022 which has its highest point at 1211399 of total actual capacity, after which it began to decrease until reaching its lowest point, 777790 in 2024.
 
 # Discussion {#sec-dis}
-## Summary Conclusions
+## Summary Conclusions and Implications
+This paper analyzes trends in shelter occupancy and capacity in Toronto from 2021 to 2024, focusing on the increasing demand for shelter services and changes in actual capacity.
+Figures above highlight that traditional shelters and motel/hotel shelters are the most commonly used services, suggesting that these facilities are critical in addressing homelessness.
+
+Furthermore, while the total service user count has generally increased, the average occupancy rates of beds and rooms have reached nearly full capacity, especially in 2024, where room occupancy approached 99.5%.
+This high occupancy rate underscores the strain on available resources and the urgent need for capacity expansion or better resource distribution.
+
+Oppose to the upward trend in shelter occupancy rates, total actual capacity met a notable decline in 2024.
+This suggests that while demand for shelter services is rising, the availability of resources has not kept pace, particularly in the latest year.
+The analysis of service user counts, occupancy rates, and capacity provides insights into the operational challenges faced by shelters in meeting the needs of the city's homeless population.
+
+## Further Discussion
+Further research should expand the time frame of analysis to understand long-term trends in shelter usage, occupancy and capacity.
+In addition, it would be beneficial to investigate the experiences and outcomes of shelter users to assess the quality and effectiveness of services provided.
+Comparative studies all over ther world facing similar homelessness challenges could provide valuable insights into better innovative solutions.
+Moreover, exploring the causes of capacity fluctuations during 2023 and 2024, such as funding changes or policy shifts, could help develop more sustainable strategies for managing shelter resources effectively.
+
+Finally, developing predictive models to forecast shelter demand could assist policymakers and social service organizations in planning and allocating resources more efficiently as the demand model can help to predict the most appropriate resources that are needed by shelter users.
+These models can incorporate not only historical data but also external factors like economic conditions, housing market trends, and social policies which can help the analysis become more comprehensive.
+
+## Limitations
+There are still some limitations in this paper.
+The data used from “Daily Shelter & Overnight Service Occupancy & Capacity” from opendatatoronto(@citeopendatatoronto) spans only four years, which may limit the ability to generalize long-term trends in shelter usage.
+Within a longer terms of trend, prediction will be more accurate for Toronto's shelter usage situation.
+In turn, seasonal variations and short-term policy changes can have significant impacts on occupancy and capacity, which might not be fully captured in this analysis.
+
+Additionally, while occupancy rates provide useful insights into shelter usage, they do not necessarily reflect the quality of services provided or the experiences of shelter users.
+Furthermore, the study relies on recorded data from the Shelter Management Information System(SMIS), which may not record all the actual situation in Toronto service programs .
+
+The consideration of external factors that can affect average occupancy rates are not included in this paper.
+As the shelter demand can change at specific times, such as during extreme weather conditions or economic downturns.
 

### 2024-09-26 scripts/01-download_data.R

@@ -12,20 +12,22 @@ library(opendatatoronto)
 library(tidyverse)
 
 #### Download data ####
-# get package
-package <- show_package("21c83b32-d5a8-4106-a54f-010dbe49f6f2")
+# Download 2021 data
+raw_data_2021 <- read_csv("https://ckan0.cf.opendata.inter.prod-toronto.ca/dataset/21c83b32-d5a8-4106-a54f-010dbe49f6f2/resource/df7d621d-a7a0-4854-81b9-8a6dc29d73a6/download/daily-shelter-overnight-service-occupancy-capacity-2021.csv")
 
-# get all resources for this package
-resources <- list_package_resources("21c83b32-d5a8-4106-a54f-010dbe49f6f2")
+# Download 2022 data
+raw_data_2022 <- read_csv("https://ckan0.cf.opendata.inter.prod-toronto.ca/dataset/21c83b32-d5a8-4106-a54f-010dbe49f6f2/resource/55d58477-50f5-490c-8da8-5953e3b26ca4/download/daily-shelter-overnight-service-occupancy-capacity-2022.csv")
 
-# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
-datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+# Download 2023 data
+raw_data_2023 <- read_csv("https://ckan0.cf.opendata.inter.prod-toronto.ca/dataset/21c83b32-d5a8-4106-a54f-010dbe49f6f2/resource/79723018-8dc7-4cd4-8fb4-8652376ea531/download/daily-shelter-overnight-service-occupancy-capacity-2023.csv")
 
-# load the first datastore resource as a sample
-data <- filter(datastore_resources, row_number()==1) %>% 
-  get_resource()
+# Download 2024 data
+raw_data_2024 <- read_csv("https://ckan0.cf.opendata.inter.prod-toronto.ca/dataset/21c83b32-d5a8-4106-a54f-010dbe49f6f2/resource/ffd20867-6e3c-4074-8427-d63810edf231/download/Daily%20shelter%20overnight%20occupancy.csv")
 
 #### Save data ####
-write_csv(data, "data/raw_data/raw_data.csv") 
+write_csv(raw_data_2021, "data/raw_data/raw_data_2021.csv")
[…]

### 2024-09-26 scripts/03-test_data.R

@@ -14,7 +14,7 @@ library(tidyverse)
 simulated_data <- read_csv("data/raw_data/simulated_data.csv")
 
 # Test for negative numbers in numerical columns
-simulated_data$SERVICE_USER_COUNT |> min() >= 0
+simulated_data$SERVICE_USER_COUNT |> min() < 0
 simulated_any_negative_bed <- any(simulated_data$CAPACITY_ACTUAL_BED |> min() < 0, 
                                   na.rm = TRUE)
 simulated_any_negative_room <- any(simulated_data$CAPACITY_ACTUAL_ROOM |> min() < 0, 
@@ -35,7 +35,7 @@ all(is.na(simulated_data$Year))
 data <- read_csv("data/analysis_data/analysis_data.csv")
 
 # Test for negative numbers in numerical columns
-data$service_user_count |> min() >= 0
+data$service_user_count |> min() < 0
 any_negative_bed <- any(data$capacity_actual_bed |> min() < 0, na.rm = TRUE)
 any_negative_room <- any(data$capacity_actual_room |> min() < 0, na.rm = TRUE)
 any_negative_rate_bed <- any(data$occupancy_rate_beds |> min() < 0, na.rm = TRUE)

### 2024-09-27 paper/paper.qmd
Final Revision

@@
 author: 
-  - "Angel Xu"
+  - "Victoria Zheng"
 thanks: "Code and data are available at: https://github.com/Anjojoo/Toronto-Overnight-Servive-Occupancy-Capacity."
@@
 
-This paper's findings are crucial for policy makers, social service organizations, and urban planners.
+This paper's findings can be helpful for policy makers, social service organizations, and urban planners.
 By shedding light on the trends in shelter occupancy and capacity, the study identifies gaps in service and potential areas for improvement.
 The analysis has implications for future resource allocation, as understanding the patterns of shelter usage can help design better support strategies, optimize capacity, and improve the overall response to homelessness in urban areas.
-Additionally, it can serve as a model for analyzing shelter systems in other cities, offering a framework for data-driven decision-making.
+Additionally, it can serve as a model for analyzing shelter systems in other cities, offering a framework for global cases.
 
 The remainder of this paper is structured as follows.
-Section 2(@sec-data) provides details on the data source, including the specific variables used, and describes the data cleaning process to prepare for analysis.
-This data section also outlines the tools and methodologies applied.
-Section 3(@sec-results) presents the key results, focusing on occupancy rates, service user counts, and actual capacity over time.
-Visualizations are used to intuitively convey these findings.
-Section 4{@sec-dis} offers a discussion of the results and their implications, summarizing conclusions and emphasizing the importance of balancing shelter occupancy and capacity for effective service provision.
+@sec-data provides details on the data source, including the specific variables used, and describes the data cleaning process to prepare for analysis.
+Data tools and methodologies applied in this paper are also stated in this section.
+Measurement and ethics are considered as well.
+@sec-results presents the key results, focusing on occupancy rates, service user counts, and actual capacity over time; Visualizations are used to intuitively convey these findings.
+@sec-dis offers a discussion of the results and their implications,also emphasizing the importance of balancing shelter occupancy and capacity for effective service provision while limitations point out some weakness in this paper.
 The comprehensive structure ensures that the reader gains a full understanding of what was analyzed, the findings, and their significance.
@@
 ## Data Measurement and Ethics
+For the data measurement of "Daily Shelter & Overnight Service Occupancy & Capacity" from opendatatoronto(@citeopendatatoronto), variables are mainly supported by Shelter Management Information System(SMIS).
+The occupancy data is recorded at 4:00 a.m.
+each morning, reflecting the previous evening's overnight service.
+Service user counts measures the number of service users at each program referring to the occupancy date.
+Actual capacity of beds and rooms are the data available on SMIS and used to calculate the occupancy rate with another available variable, occupied capacity which is also available on SMIS.
+
+The ethical consideration included ensuring that the data used reflects accurate and timely reporting from the open-source platform and respecting the privacy of the individuals indirectly represented in the data.
+The data are collected anonymized and does not contain personally identifiable information, ensuring privacy and confidentiality of shelter users.
 

### 2024-09-27 scripts/00-simulate_data.R

@@ -1,6 +1,6 @@
 #### Preamble ####
 # Purpose: Simulates the occupancy and capacity data
-# Author: Angel Xu
+# Author: Victoria Zheng
 # Date: 21 September 2024
 # Contact: victoria.zheng@mail.utoronto.ca
 # Pre-requisites: None

### 2024-09-27 scripts/01-download_data.R

@@ -1,8 +1,8 @@
 #### Preamble ####
 # Purpose: Downloads and saves the data from OpenDataToronto
-# Author: Angel Xu
+# Author: Victoria Zheng
 # Date: 24 September 2024
-# Contact: anjojoo.xu@mail.utoronto.ca
+# Contact: victoria.zheng@mail.utoronto.ca
 # Pre-requisites: None
 # Any other information needed? None
 

### 2024-09-27 scripts/02-data_cleaning.R

@@ -1,8 +1,8 @@
 #### Preamble ####
 # Purpose: Cleans the raw occupancy and capacity data recorded in the raw dataset
-# Author: Angel Xu
+# Author: Victoria Zheng
 # Date: 23 September 2024
-# Contact: anjojoo.xu@mail.utoronto.ca
+# Contact: victoria.zheng@mail.utoronto.ca
 # Pre-requisites: None
 # Any other information needed? None
 

### 2024-09-27 scripts/03-test_data.R

@@ -1,8 +1,8 @@
 #### Preamble ####
 # Purpose: Sanity check of the raw occupancy and capacity data
-# Author: Angel Xu
+# Author: Victoria Zheng
 # Date: 24 Spetember 2024
-# Contact: anjojoo.xu@mail.utoronto.ca
+# Contact: victoria.zheng@mail.utoronto.ca
 # Pre-requisites: None
 # Any other information needed? None
 

### 2024-09-27 paper/paper.qmd
Final Revision

@@
 author: 
-  - "Victoria Zheng"
+  - "Angel Xu"
 thanks: "Code and data are available at: https://github.com/Anjojoo/Toronto-Overnight-Servive-Occupancy-Capacity."

### 2024-09-27 scripts/00-simulate_data.R

@@ -1,8 +1,8 @@
 #### Preamble ####
 # Purpose: Simulates the occupancy and capacity data
-# Author: Victoria Zheng
+# Author: Angel Xu
 # Date: 21 September 2024
-# Contact: victoria.zheng@mail.utoronto.ca
+# Contact: anjojoo.xu@mail.utoronto.ca
 # Pre-requisites: None
 # Any other information needed? None
 

### 2024-09-27 scripts/01-download_data.R

@@ -1,8 +1,8 @@
 #### Preamble ####
 # Purpose: Downloads and saves the data from OpenDataToronto
-# Author: Victoria Zheng
+# Author: Angel Xu
 # Date: 24 September 2024
-# Contact: victoria.zheng@mail.utoronto.ca
+# Contact: anjojoo.xu@mail.utoronto.ca
 # Pre-requisites: None
 # Any other information needed? None
 

### 2024-09-27 scripts/02-data_cleaning.R

@@ -1,8 +1,8 @@
 #### Preamble ####
 # Purpose: Cleans the raw occupancy and capacity data recorded in the raw dataset
-# Author: Victoria Zheng
+# Author: Angel Xu
 # Date: 23 September 2024
-# Contact: victoria.zheng@mail.utoronto.ca
+# Contact: anjojoo.xu@mail.utoronto.ca
 # Pre-requisites: None
 # Any other information needed? None
 

### 2024-09-27 scripts/03-test_data.R

@@ -1,8 +1,8 @@
 #### Preamble ####
 # Purpose: Sanity check of the raw occupancy and capacity data
-# Author: Victoria Zheng
+# Author: Angel Xu
 # Date: 24 Spetember 2024
-# Contact: victoria.zheng@mail.utoronto.ca
+# Contact: anjojoo.xu@mail.utoronto.ca
 # Pre-requisites: None
 # Any other information needed? None
 

### 2024-09-27 scripts/03-test_data.R
Final Revision

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Sanity check of the raw occupancy and capacity data
 # Author: Angel Xu
-# Date: 24 Spetember 2024
+# Date: 24 September 2024
 # Contact: anjojoo.xu@mail.utoronto.ca
 # Pre-requisites: None
 # Any other information needed? None

### 2024-09-27 scripts/00-simulate_data.R
Correct data style
Applying styler for all codes

@@ -22,33 +22,42 @@ end_date <- as.Date("2024-09-01")
 number_of_datas <- 100
 
 # Define each categorical variable types
-overnight_service_type <- c("24-Hour Respite Site", "24-Hour Women's Drop-in", 
-                            "Alternative Space Protocol", 
-                            "Isolation/Recovery Site", "Motel/Hotel Shelter", 
-                            "Shelter", "Top Bunk Contingency Space", 
-                            "Warming Centre")
+overnight_service_type <- c(
+  "24-Hour Respite Site", "24-Hour Women's Drop-in",
+  "Alternative Space Protocol",
+  "Isolation/Recovery Site", "Motel/Hotel Shelter",
+  "Shelter", "Top Bunk Contingency Space",
+  "Warming Centre"
+)
 capacity_type <- c("Room Based Capacity", "Bed Based Capacity")
 
-# Simulate data 
+# Simulate data
 simulated_data <- tibble(
-  id = 1:number_of_datas, 
-  OCCUPANCY_DATE = as.Date(runif(n = number_of_datas, min = as.numeric(start_date), 
-                                        max = as.numeric(end_date)), 
-  origin = "1970-01-01"), 
-  Year = year(OCCUPANCY_DATE), 
-  OVERNIGHT_SERVICE_TYPE = sample(overnight_service_type, number_of_datas, replace = TRUE), 
-  SERVICE_USER_COUNT = rpois(n = number_of_datas, lambda = 330), 
-  CAPACITY_TYPE = sample(capacity_type, number_of_datas, replace = TRUE), 
-  CAPACITY_ACTUAL_BED = ifelse(CAPACITY_TYPE != "Bed Based Capacity", 0, 
-                               rpois(n = number_of_datas, lambda = 150)), 
-  OCCUPANCY_RATE_BEDS = ifelse(CAPACITY_TYPE != "Bed Based Capacity", 0, 
-                               sample(0:100, number_of_datas, replace = TRUE)), 
-  CAPACITY_ACTUAL_ROOM = ifelse(CAPACITY_TYPE == "Bed Based Capacity", 0, 
-                                rpois(n = number_of_datas, lambda = 170)), 
-  OCCUPANCY_RATE_ROOMS = ifelse(CAPACITY_TYPE != "Bed Based Capacity", 0, 
[…]

### 2024-09-27 scripts/01-download_data.R

@@ -29,5 +29,3 @@ write_csv(raw_data_2021, "data/raw_data/raw_data_2021.csv")
 write_csv(raw_data_2022, "data/raw_data/raw_data_2022.csv")
 write_csv(raw_data_2023, "data/raw_data/raw_data_2023.csv")
 write_csv(raw_data_2024, "data/raw_data/raw_data_2024.csv")
-
-

### 2024-09-27 scripts/02-data_cleaning.R

@@ -23,7 +23,7 @@ raw_data_2024 <- read_csv("data/raw_data/raw_data_2024.csv")
 # Convert occupancy_date into date format
 raw_data_2021$OCCUPANCY_DATE <- as.Date(raw_data_2021$OCCUPANCY_DATE, format = "%y-%m-%d")
 raw_data_2022$OCCUPANCY_DATE <- as.Date(raw_data_2022$OCCUPANCY_DATE, format = "%y-%m-%d")
-raw_data_2023$OCCUPANCY_DATE<- as.Date(raw_data_2023$OCCUPANCY_DATE, format="%Y-%m-%dT%H:%M:%S")
+raw_data_2023$OCCUPANCY_DATE <- as.Date(raw_data_2023$OCCUPANCY_DATE, format = "%Y-%m-%dT%H:%M:%S")
 raw_data_2024$OCCUPANCY_DATE <- as.Date(raw_data_2024$OCCUPANCY_DATE, format = "%y-%m-%d")
 
 # Combine 2021-2024 datasets into one dataset
@@ -32,17 +32,19 @@ raw_data <- rbind(raw_data_2021, raw_data_2022, raw_data_2023, raw_data_2024)
 # Clean, mutate and select wanted columns
 cleaned_data <-
   raw_data |>
-  janitor::clean_names() |> 
-  mutate(year = year(occupancy_date)) |> 
-  select(year, 
-         occupancy_date, 
-         overnight_service_type, 
-         service_user_count, 
-         capacity_type, 
-         capacity_actual_bed,
-         occupancy_rate_beds,
-         capacity_actual_room,
-         occupancy_rate_rooms) |> 
+  janitor::clean_names() |>
+  mutate(year = year(occupancy_date)) |>
+  select(
+    year,
+    occupancy_date,
+    overnight_service_type,
+    service_user_count,
+    capacity_type,
+    capacity_actual_bed,
+    occupancy_rate_beds,
+    capacity_actual_room,
+    occupancy_rate_rooms
+  ) |>
   drop_na(overnight_service_type)
 
 # View the cleaned dataset (optional)

### 2024-09-27 scripts/03-test_data.R

@@ -15,16 +15,22 @@ simulated_data <- read_csv("data/raw_data/simulated_data.csv")
 
 # Test for negative numbers in numerical columns
 simulated_data$SERVICE_USER_COUNT |> min() < 0
-simulated_any_negative_bed <- any(simulated_data$CAPACITY_ACTUAL_BED |> min() < 0, 
-                                  na.rm = TRUE)
-simulated_any_negative_room <- any(simulated_data$CAPACITY_ACTUAL_ROOM |> min() < 0, 
-                                   na.rm = TRUE)
-simulated_any_negative_rate_bed <- any(simulated_data$OCCUPANCY_RATE_BEDS |> min() < 0, 
-                                       na.rm = TRUE)
-simulated_any_negative_rate_room <- any(simulated_data$OCCUPANCY_RATE_ROOMS |> min() < 0, 
-                                        na.rm = TRUE)
-all(simulated_any_negative_bed, simulated_any_negative_room, 
-    simulated_any_negative_rate_bed, simulated_any_negative_rate_room)
+simulated_any_negative_bed <- any(simulated_data$CAPACITY_ACTUAL_BED |> min() < 0,
+  na.rm = TRUE
+)
+simulated_any_negative_room <- any(simulated_data$CAPACITY_ACTUAL_ROOM |> min() < 0,
+  na.rm = TRUE
+)
+simulated_any_negative_rate_bed <- any(simulated_data$OCCUPANCY_RATE_BEDS |> min() < 0,
+  na.rm = TRUE
+)
+simulated_any_negative_rate_room <- any(simulated_data$OCCUPANCY_RATE_ROOMS |> min() < 0,
+  na.rm = TRUE
+)
+all(
+  simulated_any_negative_bed, simulated_any_negative_room,
+  simulated_any_negative_rate_bed, simulated_any_negative_rate_room
+)
 
 # Test for NA in categorical columns
 all(is.na(simulated_data$OVERNIGHT_SERVICE_TYPE))
@@ -46,5 +52,3 @@ all(any_negative_bed, any_negative_room, any_negative_rate_bed, any_negative_rat
 all(is.na(data$overnight_service_type))
 all(is.na(data$capacity_type))
 all(is.na(data$year))
-
-

### 2024-09-27 paper/paper.qmd
Final Revision
Correct data style
Applying styler for all codes
Correct code styles for the paper

@@
 Tidyverse[@citetidyverse], here[@citehere], ggplot2[@citeggplot2], dplyr[@citedplyr], lubridate[@citelubridate], janitor[@citejanitor] and knitr[@citeknitr] are all applied in the analysis.
+Styler[@citestyler] is used to correct code styles.
 
@@
 #  Create a subset includes key variables
-subset <- data[, c("year", "overnight_service_type", "service_user_count", "occupancy_rate_beds", 
-                   "occupancy_rate_rooms")]
+subset <- data[, c(
+  "year", "overnight_service_type", "service_user_count", "occupancy_rate_beds",
+  "occupancy_rate_rooms"
+)]
 
@@
 # Present the sample as a table
-kable(sample, col.names = c("year", "overnight service type", "service user count", 
-                            "occupancy rate beds", "occupancy rate rooms"), 
+kable(sample,
+  col.names = c(
+    "year", "overnight service type", "service user count",
+    "occupancy rate beds", "occupancy rate rooms"
+  ),
   align = "ccc",
-      caption = "Table 1: Sample of the key variables in Occupancy & Capacity Data")
+  caption = "Table 1: Sample of the key variables in Occupancy & Capacity Data"
+)
 ```
@@
 #  Create a subset includes key variables
-summary_stats <- data.frame(Variable = c("Service User Count", "Occupancy Rate of Beds", 
-                                         "Occupancy Rate of Rooms"), 
+summary_stats <- data.frame(
+  Variable = c(
+    "Service User Count", "Occupancy Rate of Beds",
+    "Occupancy Rate of Rooms"
+  ),
   Mean = c(mean_user_cnt, mean_rate_bed, mean_rate_room),
-                            SD = c(sd_user_cnt, sd_rate_bed, sd_rate_room))
+  SD = c(sd_user_cnt, sd_rate_bed, sd_rate_room)
+)
 
 # Present the sample as a table
-kable(summary_stats, col.names = c("Variable", "Mean", "Standard Deviation"), 
+kable(summary_stats,
+  col.names = c("Variable", "Mean", "Standard Deviation"),
   align = "ccc",
-      caption = "Table 2: Mean and Standard Deviation of Key Variables")
+  caption = "Table 2: Mean and Standard Deviation of Key Variables"
+)
 ```
@@
 #  Create a subset includes key variables
-summary_stats <- data.frame(Year = c("2021", "2022", "2023", "2024"), 
-                            Service_User_Counts_Sum = c(sum_cnt_2021, sum_cnt_2022, 
-                                                       sum_cnt_2023, sum_cnt_2024), 
-                            Actual_Bed_Capacity_Sum = c(sum_bed_2021, sum_bed_2022, 
-                                                        sum_bed_2023, sum_bed_2024), 
-                            Actual_Room_Capacity_Sum = c(sum_room_2021, sum_room_2022, 
-                                                         sum_room_2023, sum_room_2024)
+summary_stats <- data.frame(
+  Year = c("2021", "2022", "2023", "2024"),
+  Service_User_Counts_Sum = c(
+    sum_cnt_2021, sum_cnt_2022,
+    sum_cnt_2023, sum_cnt_2024
+  ),
+  Actual_Bed_Capacity_Sum = c(
+    sum_bed_2021, sum_bed_2022,
+    sum_bed_2023, sum_bed_2024
+  ),
+  Actual_Room_Capacity_Sum = c(
+    sum_room_2021, sum_room_2022,
+    sum_room_2023, sum_room_2024
+  )
 )
@@
 # Present the sample as a table
-kable(summary_stats, col.names = c("Year", "Total Service User Counts", 
-                                   "Total Actual Bed Capacity", "Total Actual Room Capacity"), 
+kable(summary_stats,
+  col.names = c(
+    "Year", "Total Service User Counts",
+    "Total Actual Bed Capacity", "Total Actual Room Capacity"
+  ),
   align = "ccc",
-      caption = "Table 3: Total Service User Count and Actual Capacity between years")
+  caption = "Table 3: Total Service User Count and Actual Capacity between years"
+)
 ```
@@
   scale_y_continuous(labels = scales::comma)
-
 ```
@@
 # Reshape the data to long format for plotting with side-by-side bars
-df_long <- tidyr::pivot_longer(avg_rate, cols = c("avg_rate_bed", "avg_rate_room"),
+df_long <- tidyr::pivot_longer(avg_rate,
+  cols = c("avg_rate_bed", "avg_rate_room"),
   names_to = "type",
-                               values_to = "average")
+  values_to = "average"
+)
 
@@
     position = position_dodge(width = 0.9),
-            vjust = -0.5, size = 3.5)
-
+    vjust = -0.5, size = 3.5
+  )
 ```
@@
 # Reshape the data to long format for plotting with side-by-side bars
-df_long <- tidyr::pivot_longer(total_cap, cols = c("total_cap_bed", "total_cap_room"),
+df_long <- tidyr::pivot_longer(total_cap,
+  cols = c("total_cap_bed", "total_cap_room"),
   names_to = "type",
-                               values_to = "total")
+  values_to = "total"
+)
 
@@
   theme_minimal()
-
 ```

### 2024-09-27 paper/paper.qmd
Final revision of readme and paper
Add data cleaning details in the paper, change data files explanation in readme

@@
 @sec-dis offers a discussion of the results and their implications,also emphasizing the importance of balancing shelter occupancy and capacity for effective service provision while limitations point out some weakness in this paper.
+@sec-app provides additional details for the paper, including process of data cleaning.
 The comprehensive structure ensures that the reader gains a full understanding of what was analyzed, the findings, and their significance.
@@
 Prior to analysis, data cleaning is essential to remove unnecessary columns and create new ones for further examination. 
-The occupancy date variable is converted to a date format, allowing the extraction of the year for subsequent analysis.
-After this conversion, a new 'year' column is mutated to show trends in service user counts and occupancy rates from 2021 to 2024.
-Once these steps are complete, relevant variables are selected from the raw dataset to simplify and clarify the analysis.
+New "year" column is mutated to show trends in service user counts and occupancy rates from 2021 to 2024.
+Afterwards, relevant variables are selected from the raw dataset to simplify and clarify the analysis.
 The chosen variables include year, occupancy date, overnight service type, service user count, capacity type, actual capacity and occupancy rates.
@@
 
+# Appendix {#sec-appendix}
+## Details of Data Cleaning
+Occupancy date of 2021-2024 raw data are all transformed into date format for following new column of "year" mutation which represents the year of occupancy and capacity data that are collected.
+After the transformation, a combination work of these four raw dataset is done for more efficient following variables selection among all data.
+NAs are cleaned in categorical variable of overnight service type.
+
 \newpage

### 2024-09-27 paper/paper.qmd
Final Revision of paper

@@
 
-# Appendix {#sec-appendix}
+# Appendix {#sec-app}
 ## Details of Data Cleaning
